Statistical Theory\\ Chess Dataset Analysis

### 2024-09-29 Article/Article.tex
Updated Article

Our dataset \cite{dataset} comprises 19,113 games from Lichess, an online chess platform, each described by 16 features. The most relevant features to our analysis are: \begin{itemize} \item \textbf{Turns}: A positive integer representing the length of the game, where one turn includes a move by both White and Black. \item \textbf{ELO Rating}: A system that quantifies player skill, widely used by FIDE and online chess platforms \cite{ELO}, which we use to assess its correlation with match length. \item \textbf{Time Controls}: Competitive games are categorized into four main time control classes: Bullet, Blitz, Rapid, and Classical, based on initial clock times and increments. \item \textbf{Winner and Victory Status}: These columns capture whether White or Black won and the method of victory (checkmate, resignation, draw). \end{itemize} Through statistical analysis, we aim to understand how these factors influence the duration of chess matches and to provide insights into the relationship between skill level and game dynamics.

@@

\subsubsection{Kolmogorov-Smirnov (KS) Test for Normality}
We applied the Kolmogorov-Smirnov (KS) test to evaluate whether the player rating distribution follows a normal distribution. The test returned a p-value of \(8.9649 \times 10^{-13}\), which is significantly lower than the threshold of \(0.03\) for a confidence level of 97\%, thus rejecting the null hypothesis that the rating distribution is normally distributed. However, when the test was performed on a random 10\% sample of player ratings, the p-value increased to 0.49, meaning we could not reject the null hypothesis at that sample size. This result indicates that while the entire population does not fit a normal distribution, smaller samples may approximate normality.

>>>>

Our dataset \cite{dataset} comprises 19,113 games from Lichess, an online chess platform, each described by 16 features. The most relevant features to our analysis are: \begin{itemize} \item \textbf{Turns}: A positive integer representing the length of the game, where one turn includes a move by both White and Black. \item \textbf{ELO Rating}: A system that quantifies player skill, widely used by FIDE and online chess platforms \cite{ELO}, which we use to assess its correlation with match length. \item \textbf{Time Controls}: Competitive games are categorized into four main time control classes: Bullet, Blitz, Rapid, and Classical, based on initial clock times and increments \cite{time}. \item \textbf{Winner and Victory Status}: These columns capture whether White or Black won and the method of victory (checkmate, resignation, draw). \end{itemize} Through statistical analysis, we aim to understand how these factors influence the duration of chess matches and to provide insights into the relationship between skill level and game dynamics.

@@

\subsubsection{Kolmogorov-Smirnov (KS) Test for Normality}
We applied the Kolmogorov-Smirnov (KS) test to evaluate whether the player rating distribution follows a normal distribution. The test returned a p-value of \(8.9649 \times 10^{-13}\), which is significantly lower than the threshold of \(0.03\) for a confidence level of 97\%, thus rejecting the null hypothesis that the rating distribution is normally distributed. However, when the test was performed on a random 10\% sample of player ratings, the p-value increased to 0.49, meaning we could not reject the null hypothesis at that sample size. This result indicates that while the entire population does not fit a normal distribution, smaller samples may approximate normality \cite{kssample}.
